---
title: Synthetic Assessment Data Pipeline (SAD-Pipe) - Simulation
output: html_document
editor_options: 
  markdown: 
    wrap: 120
---

This notebook generates a synthetic multiple-choice exam dataset with realistic psychometric structure.

## Purpose

The goal is to produce *reproducible* synthetic response data for prototyping and testing downstream analysis workflows
(e.g., scoring, item analysis, distractor behavior).

The simulation combines:

-   2PL-like knowledge model (item difficulty `b`, discrimination `a`)
-   person-level traits (`theta_true`, `carelessness`, `guessiness`, `testwiseness`)
-   item types (regular / regular_clean / diluted / misleading)
-   distractor attractiveness profiles

## Reproducibility

-   Randomness is controlled via `seed` .
-   With the same `seed` and the same `simulation_config`, the script is intended to generate identical outputs.

## Configuration

The main entry point is:

-   `results = run_simulation(config)`

Key parameters:

-   `seed`: controls randomness
-   `n_items`: number of items
-   `n_students`: number of test-takers
-   `options`: answer options (default: `A`–`E`)
-   `item_type_dist`: item type proportions

### Item type proportions

Set proportions via `item_type_dist` in `simulation_config`. You can omit `regular`; it will be set to the remainder.

``` r
config = simulation_config(
    seed=42,
    n_items=80,
    n_students=200,
    item_type_dist={
        "regular_clean": 0.20, # 20% of the items
        "diluted": 0.10, # 10 % of the items
        "misleading": 0.10, # 10 % of the items
        # "regular" becomes 0.600 automatically
    },
)
results = run_simulation(config)
```

## Outputs (data dictionary)

The simulation returns a dictionary `results` with four data frames (tibbles).

### `results$answer_matrix`

-   **Shape**: `(n_students, n_items)`
-   **Rows**: one row per person
-   **Columns**: `item1 ... itemN`
-   **Values**: selected option label (e.g., `"A"`, `"B"`, ...)

### `results$scored_responses`

-   **Item columns**: `0/1` correctness per item
-   **`score`**: sum over item columns
-   **`score_pct`**: `score / n_items * 100`

### `results$item_metadata`

One row per item.

-   **`item`**: item identifier (e.g., `item17`)
-   **`item_type`**: one of `regular`, `regular_clean`, `diluted`, `misleading`
-   **`num_options`**: number of answer options (including the correct answer)
-   **`correct_answer`**: correct option label
-   **`a_param`**: discrimination
-   **`b_param`**: difficulty
-   **`slip`**: slip probability (know → incorrect)
-   **`distractors`**: list of incorrect option labels
-   **`distractor_weights`**: mapping distractor → sampling weight

### `results$testtakers`

One row per person.

-   **`theta_true`**: latent ability
-   **`carelessness`**, **`guessiness`**, **`testwiseness`**: behavioral traits
-   **`score`**, **`score_pct`**: total performance summary

## Code and Preparation

```{r library}
#| include: false
#| warning: false

library(tidyverse)
library(fs)
library(rlang)
library(patchwork)
library(ggplot2)
```

```{r plotthemes}
#| include: false

# Seaborn Theme in ggplot2 (whitegrid, pastel)
theme_set(
  theme_minimal() + 
  theme(
    panel.grid.major = element_line(color = "#e0e0e0"),
    panel.grid.minor = element_blank()
  )
)

# Farbpalette 
item_type_colors <- c(
  "regular"       = "#3B5070",
  "regular_clean" = "#765B7A",
  "diluted"       = "#B56576",
  "misleading"    = "#eaac8b"
)
```

### Simulation Configuration Function

```{r simulation_config_function}

simulation_config <- function(
    seed = 63111426,
    n_items = 80,
    n_students = 200,
    options = c("A", "B", "C", "D", "E"),
    output_dir = "../data",
    item_type_dist = NULL,
    theta = NULL
) {
  # setting expected item type keys 
  expected <- c("regular", "regular_clean", "diluted", "misleading")
  
  if (is.null(item_type_dist)) {
    dist <- list(
      regular_clean = 0.20,
      diluted       = 0.10,
      misleading    = 0.10,
      regular       = 0.60
    )
  } else {

    unknown <- setdiff(names(item_type_dist), expected)
    if (length(unknown) > 0) {
      stop(paste("item_type_dist contains unknown keys:", paste(sort(unknown), collapse = ", ")))
    }
    
    # 
    if (any(unlist(item_type_dist) < 0)) {
      stop("item_type_dist contains negative values")
    }
    
    dist <- item_type_dist
    total_pre <- sum(unlist(dist))
    
    if (total_pre < 0) {
      stop("item_type_dist must sum to a positive number")
    }
    
    if (total_pre > 1) {
      stop(paste("Item type probabilities sum to", total, "; expected <= 1.0"))
    }
    
    # set regular to rest or missing probalities so that sum of all type probabilities is 1
    if (!"regular" %in% names(dist)) {
      dist$regular <- 1 - total_pre
    }
    
    total <- sum(unlist(dist))
    
    if (!isTRUE(all.equal(total, 1))) {
      dist$regular <- dist$regular + (1 - total)
    }

    # final sum validation
    if (!isTRUE(all.equal(sum(unlist(dist)), 1.0))) {
      stop("Item type probabilities sum to ", sum(unlist(dist)), ", expected 1.0")
    }
  }

    if (!is.null(theta)) {
      if(length(theta) != n_students) {
        stop("Length of theta and intented number of students do not match. Please check.")
      }
    }

  # after checks, return correct objects
  structure(
    list(
      seed = seed,
      n_items = n_items,
      n_students = n_students,
      options = options,
      output_dir = output_dir,
      theta = theta,
      item_type_dist = dist[expected]
    ),
    class = "SimulationConfig"
  )
}


```

example call:

```{r examplesim_conf_call}
config <- simulation_config(n_items = 100)
print(config$item_type_dist)
```

### Item Generator Function

```{r item_generator}

item_generator <- function(config) {
  set.seed(config$seed)

  # Parameter spezifications
  item_param_specs <- list(
    "regular_clean" = list(b_loc = -0.3, b_scale = 0.6, a_low = 0.8, a_high = 1.7),
    "regular"       = list(b_loc = -0.7, b_scale = 1.2, a_low = 0.5, a_high = 1.6),
    "diluted"       = list(b_loc = -1.1, b_scale = 0.8, a_low = 0.1, a_high = 0.7),
    "misleading"    = list(b_loc = -0.2, b_scale = 0.9, a_low = 0.4, a_high = 1.7)
  )

  distractor_specs <- list(
    "regular_clean" = list(prob = 0.0, bump = 0.0, force_dominant = FALSE),
    "regular"       = list(prob = 0.3, bump = 0.15, force_dominant = FALSE),
    "diluted"       = list(prob = 0.7, bump = 0.30, force_dominant = FALSE),
    "misleading"    = list(prob = 0.95, bump = 0.60, force_dominant = TRUE)
  )

  # Internal support function: Assign item types (largest remainder method)
  assign_item_types <- function() {
    n <- config$n_items
    keys <- c("regular_clean", "diluted", "misleading", "regular")
    
    raw <- sapply(keys, function(k) n * config$item_type_dist[[k]])

    base <- floor(raw)
    remainder <- n - sum(base)
    
    if (remainder > 0) {
      diffs <- sort(raw - base, decreasing = TRUE)
      to_add <- names(diffs)[1:remainder]
      for (k in to_add) base[k] <- base[k] + 1
    }
    types <- rep(names(base), base)
    return(sample(types)) # Shuffle
  }

  # Internal support function: distractor weights
  generate_distractor_weights <- function(distractors, item_type) {
    k <- length(distractors)
    if (k == 0) return(numeric(0))

    weights <- if (k == 1) {
      1.0
    } else if (k == 2) {
      c(0.6, 0.4)
    } else {
      c(0.45, 0.30, rep(0.25 / (k - 2), k - 2))
    }
    weights <- weights / sum(weights)

    spec <- distractor_specs[[item_type]]
    if (runif(1) < spec$prob && k > 1) {
      idx <- sample(1:k, 1)
      weights[idx] <- weights[idx] + spec$bump
      weights <- weights / sum(weights)
      
      if (spec$force_dominant) {
        dom_idx <- which.max(weights)
        weights[dom_idx] <- max(weights[dom_idx], 0.6)
        weights <- weights / sum(weights)
      }
    }
    names(weights) <- distractors
    return(list(weights))
  }

  # Main function: generate items
  generate_items <- function() {
    item_types <- assign_item_types()
    items <- list()
    answer_key <- list()
    
    for (i in seq_along(item_types)) {
      type <- item_types[i]
      id <- paste0("item", i)
      
      num_opts <- sample(c(3, 4, 5), 1, prob = c(0.25, 0.45, 0.30))
      opts <- config$options[1:num_opts]
      correct <- sample(opts, 1)
      distractors <- setdiff(opts, correct)

      spec <- item_param_specs[[type]]
      item <- list(
        item = id,
        item_type = type,
        num_options = num_opts,
        correct_answer = correct,
        a_param = runif(1, spec$a_low, spec$a_high),
        b_param = rnorm(1, spec$b_loc, spec$b_scale),
        slip = rbeta(1, 1.5, 30),
        distractors = list(distractors),
        distractor_weights = generate_distractor_weights(distractors, type)
      )
      items[[id]] <- item
      answer_key[[id]] <- correct
    }
    return(list(items = items, answer_key = answer_key))
  }

  return(list(generate_items = generate_items))
}
```

### Response Simulator Function

```{r response_simulator}
response_simulator <- function(config) {
  set.seed(config$seed)

  # Main function
  simulate_responses <- function(items, n_students) {
  # Generate abilities and characteristics
    data <- generate_student_abilities(config, n_students)
    theta <- data$theta
    traits <- data$traits
    n_items <- length(items)
    item_ids <- sapply(items, function(x) x$item)
    answer_mat <- matrix(NA, nrow = n_students, ncol = n_items, 
                         dimnames = list(NULL, item_ids))
    score_mat  <- matrix(NA, nrow = n_students, ncol = n_items, 
                         dimnames = list(NULL, item_ids))

    for (i in seq_along(items)) {
      item <- items[[i]]
      item_id <- item$item
      responses <- simulate_item_responses(item, theta, traits)
      answer_mat[, item_id] <- responses$answers
      score_mat[, item_id]  <- responses$scores
    }

    scored_responses <- as_tibble(score_mat) %>%
      mutate(
        score = rowSums(score_mat),
        score_pct = round((score / n_items) * 100, 2)
      )

    testtakers <- as_tibble(traits) %>%
      mutate(theta_true = theta) %>%
      relocate(theta_true) %>% 
      bind_cols(scored_responses %>% select(score, score_pct))

    return(list(
      answer_matrix = as_tibble(answer_mat),
      scored_responses = scored_responses,
      testtakers = testtakers
      ))
  }

  # Internal support function
  generate_student_abilities <- function(config, n_students) {
    # Ability (Theta)
    if (!is.null(config$theta)) {
      theta <- as.numeric(config$theta)
      if (length(theta) != n_students) stop("Provided theta array must match n_students.")
    } else {
      # theta default values
      theta_main <- rbeta(n_students, 8, 2) * 6 - 4
      theta_elite <- rnorm(n_students, 2.6, 0.4)
      mix <- runif(n_students) < 0.03
      theta <- ifelse(mix, theta_elite, theta_main)
    }
  
    # Top performers Identification
    top_mask <- theta > quantile(theta, 0.88)
  
    # Behavioral Traits
    traits <- list(
      carelessness = pmin(pmax(rbeta(n_students, 2, 12) * 0.10, 0.0), 0.08),
      guessiness   = pmin(pmax(rbeta(n_students, 2, 9) * 0.40, 0.00), 0.28),
      testwiseness = rnorm(n_students, -0.20, 0.20)
    )

    # Ability-dependent testwiseness
    base_tw <- rnorm(n_students, 0.10, 0.20)
    severity <- pmax(1.0 - theta, 0.0)
    if (max(severity) > 0) {
      severity <- 0.7 * (severity / max(severity)) + 0.3
    }
    traits$testwiseness <- base_tw - severity

    # Adjust traits for top performers
    traits$carelessness[top_mask] <- traits$carelessness[top_mask] * 0.85
    #traits$guessiness[top_mask]   <- traits$guessiness[top_mask] * 0.85

    return(list(theta = theta, traits = traits))
  }

  # Internal support function
  simulate_item_responses <- function(item, theta, traits) {
    n_students <- length(theta)
    a <- item$a_param
    b <- item$b_param
    slip <- item$slip
    correct <- item$correct_answer
    distractors <- item$distractors[[1]]
    weights <- item$distractor_weights[[1]]

    answers <- character(n_students)
    scores <- numeric(n_students)

    p_know <- 1 / (1 + exp(-a * (theta - b)))
    p_know <- pmin(pmax(p_know + rnorm(n_students, 0, 0.05), 0), 1)
  
    # Special handling misleading
    if (item$item_type == "misleading") {
      theta_adj <- theta + rnorm(n_students, 0, 0.4)
      p_know <- 1 / (1 + exp(-a * (theta_adj - b)))
    }

    dominant_distractor <- if (length(weights) > 0) names(which.max(weights)) else distractors[1]

    for (i in 1:n_students) {
      knows <- runif(1) < p_know[i]
      
      # Carelessness
      if (knows && runif(1) < traits$carelessness[i]) {
        answers[i] <- choose_incorrect(correct, distractors, weights)
        scores[i] <- 0
        next
      }

      # Misleading Trap
      if (item$item_type == "misleading") {
        in_trap_band <- theta[i] < -0.5
        if (in_trap_band && !knows) {
          if (runif(1) < 0.85) {
            answers[i] <- dominant_distractor
            scores[i] <- 0
            next
          }
        }
      }

      # Knows answer (incl. slip)
      if (knows) {
        if (runif(1) < slip) {
          answers[i] <- choose_incorrect(correct, distractors, weights)
          scores[i] <- 0
        } else {
          answers[i] <- correct
          scores[i] <- 1
        }
        next
      }
  
      # Does not know (Guessing / Testwiseness)
      # Testwiseness (Middle option bias)
      if (traits$testwiseness[i] < -0.05 && length(distractors) > 1) {
        middle_option <- distractors[floor(length(distractors)/2) + 1]
        if (runif(1) < 0.60) {
          answers[i] <- middle_option
          scores[i] <- ifelse(answers[i] == correct, 1, 0)
          next
        }
      }

      # Guessiness bias
      if (runif(1) < traits$guessiness[i]) {
        n_options <- length(distractors) + 1
        if (runif(1) < (1/n_options)) {
          answers[i] <- correct
        } else {
          answers[i] <- sample(distractors, 1)
        }
        scores[i] <- ifelse(answers[i] == correct, 1, 0)
        next
      }

      # Fallback
      all_opts <- c(correct, distractors)
      answers[i] <- sample(all_opts, 1)
      scores[i] <- ifelse(answers[i] == correct, 1, 0)
    }

    return(list(answers = answers, scores = scores))
  }

  # Internal support function
  choose_incorrect <- function(correct, distractors, weights) {
    if (length(distractors) == 0) return(correct)
    if (length(weights) == 0) return(sample(distractors, 1))

    return(sample(names(weights), 1, prob = weights))
  }

  return(list(simulate_responses = simulate_responses))
}
```

### Main Function: Run Simulation

```{r run_simulation}
run_simulation <- function(config = NULL) {

  if(is.null(config)){
    config <- simulation_config() 
  }

  if (!dir.exists(config$output_dir)) {
    dir.create(config$output_dir, recursive = TRUE)
  }
  
  # initialize components
  item_gen <- item_generator(config)
  resp_sim <- response_simulator(config)
  
  message("Generating test items...")
  gen_data <- item_gen$generate_items()
  items <- gen_data$items
  item_metadata <- bind_rows(items)

  message(paste("Simulating responses for", config$n_students, "students..."))
  sim_output <- resp_sim$simulate_responses(items, config$n_students)
  
  results <- list(
    answer_matrix    = sim_output$answer_matrix,
    item_metadata    = item_metadata,
    scored_responses = sim_output$scored_responses,
    testtakers       = sim_output$testtakers
  )

  message(paste("Saving results to", config$output_dir, "..."))

  for (name in names(results)) {
    df <- results[[name]]
    base_path <- file.path(config$output_dir, name)
    
    write_csv(df, paste0(base_path, "_R.csv"))
    saveRDS(df, paste0(base_path, ".rds"))
  }

  class(results) <- "SimulationResults"

  return(results)
}

```

## Step 0 — Configure Your Test

In this block you define the structure of the exam itself:

-   total number of items,
-   proportions of each item type,
-   and the remaining proportion will automatically become “regular” items.

Feel free to adjust these values to create easier, harder, or more diagnostic tests.

### Test Blueprint Parameters

```{r default_values}
n_items <- 80   # Total number of items in the exam

# Proportions of special item types
regular_clean <- 0.30   # High-quality items with clean distractors
misleading    <- 0.10   # Items with attractive traps or cues
diluted       <- 0.10   # Weak or unfocused items with low discrimination

# The remaining proportion (1 - sum above) becomes "regular" items automatically
```

## Step 1 — Construct Your Student Ability Distribution (θ)

This block lets you design and preview the ability distribution for your simulated student population.

You can:

-   adjust the Beta parameters for the main population,
-   control the proportion and strength of the elite tail,
-   inspect the resulting θ distribution visually,
-   and decide whether to use it for the simulation.

Feel free to adjust the parameters until the curve looks realistic for your scenario. When you're satisfied, the
generated `theta_preview` will be passed into the main simulation.

If you prefer the default distribution, simply run the cell without changes.

These are the base parameters:

***Main population (right-skewed)***

-   beta_a = 7
-   beta_b = 2
-   scale = 4
-   shift = -3.2

***Elite tail***

-   elite_mean = 2.5
-   elite_sd = 1
-   elite_fraction = 0.1

```{r default_values2}
#| results: 'asis'
#| echo: false
# --- Ability Construction & Preview ---

# Choose your parameters here
seed <- 63111426
#seed <- 42
n_students <- 200


# Main population (right-skewed Beta distribution)
beta_a <- 7      # Shape parameter α: higher values push mass toward the right (more good students)
beta_b <- 2      # Shape parameter β: lower values keep the distribution skewed (few low-ability students)
scale <- 4       # Stretch factor: controls the overall width of the ability range
shift <- -3.2    # Shift factor: moves the center of the distribution left/right on the θ scale

# Elite tail (small group of very strong students)
elite_mean <- 2.5       # Average ability level of elite students (controls how "excellent" they are)
elite_sd <- 1            # Spread of elite abilities (tight cluster vs. more variation)
elite_fraction <- 0.1  # Proportion of students drawn from the elite tail (e.g., 3% very strong students)


# Generate
set.seed(seed)
theta_main  <- rbeta(n_students, beta_a, beta_b) * scale + shift
theta_elite <- rnorm(n_students, mean = elite_mean, sd = elite_sd)
mix         <- runif(n_students) < elite_fraction

theta_preview <- ifelse(mix, theta_elite, theta_main)

pl <- ggplot(data.frame(theta = theta_preview), aes(x = theta)) +
  geom_density(fill = "#b5c7d3", alpha = 0.5) + # Pastel-Farbe ähnlich Seaborn
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray") +
  labs(
    title = "Preview: Ability Distribution (θ)",
    x = "θ",
    y = "Density"
  ) +
  theme_minimal()

show(pl)

cat(sprintf("\n\nmin=%.2f, max=%.2f, mean=%.2f\n", 
            min(theta_preview), max(theta_preview), mean(theta_preview)))

cat(sprintf("\nStudents above 2.5: %d\n", sum(theta_preview > 2.5)))

```

## Step 2 — Run the Simulation

If you're happy with the θ distribution above, pass `theta_preview` into the simulation via
`simulation_config(theta=theta_preview)`.

If not, go back to Step 1, adjust the parameters, and preview again.

```{r config_run}
# Config and run the simulation
config <- simulation_config(
    seed = seed,
    n_items = n_items,
    n_students = n_students,
    output_dir = "../data/simulation_results",
    item_type_dist = list(
        "regular_clean" = regular_clean,
        "misleading"    = misleading,
        "diluted"       = diluted
    ),
   theta = theta_preview
)

results <- run_simulation(config)
```


```{r run_show}
#| echo: false
cat("Sample of generated data:")
cat("Answer matrix (first 5 students, first 5 items):")

print(results$answer_matrix[1:5, 1:5])

cat("Item metadata (first 5 items):")

print(head(results$item_metadata, 5))

cat("Test-taker scores (first 5 students):")

print(head(results$testtakers[, c("theta_true", "carelessness", "score")], 5))
```

```{r summary_statistics}
#| echo: false


print_summary_statistics <- function(results) {
  testtakers <- results$testtakers
  item_metadata <- results$item_metadata
  scored_responses <- results$scored_responses
  
  cat("=== Test-Taker Summary ===\n")
  cat(sprintf("Number of students: %d\n", nrow(testtakers)))
  cat(sprintf("Mean score (%%): %.1f ± %.1f\n", 
              mean(testtakers$score_pct), sd(testtakers$score_pct)))
  cat(sprintf("Score range (%%): %.1f - %.1f\n", 
              min(testtakers$score_pct), max(testtakers$score_pct)))
  cat(sprintf("Mean ability (theta): %.3f ± %.3f\n", 
              mean(testtakers$theta_true), sd(testtakers$theta_true)))
  
  cat("\n=== Item Analysis ===\n")
  cat(sprintf("Number of items: %d\n", nrow(item_metadata)))
  cat(sprintf("Mean difficulty (b): %.3f ± %.3f\n", 
              mean(item_metadata$b_param), sd(item_metadata$b_param)))
  cat(sprintf("Mean discrimination (a): %.3f ± %.3f\n", 
              mean(item_metadata$a_param), sd(item_metadata$a_param)))
  
  cat("\nItem Type Distribution:\n")
  # table() entspricht value_counts()
  print(table(item_metadata$item_type))
  
  # Item-Total Korrelation berechnen
  # Wir entfernen die Summen-Spalten (score, score_pct)
  item_scores <- scored_responses %>% select(-score, -score_pct)
  total_scores <- scored_responses$score
  
  # cor() berechnet die Korrelation zwischen Matrix-Spalten und einem Vektor
  item_total_corrs <- cor(item_scores, total_scores)
  
  cat(sprintf("\nMean item-total correlation: %.3f\n", mean(item_total_corrs)))
}

# Aufruf der Funktion
print_summary_statistics(results)

```

```{r plots1}

# 1. Distribution of Item Types
plot_item_type_distribution <- function(item_metadata) {
  ggplot(item_metadata, aes(x = item_type, fill = item_type)) +
    geom_bar() +
    scale_fill_manual(values = item_type_colors) +
    labs(title = "Distribution of Item Types", x = "Item Type", y = "Count") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1),
          legend.position = "none",
          panel.grid.major.x = element_blank())
}

# 2. Answer Options and Correct Answers
plot_answer_options <- function(item_metadata) {
  # Falls num_options fehlt, berechnen (Länge distractors + 1)
  if (!"num_options" %in% names(item_metadata)) {
    item_metadata$num_options <- sapply(item_metadata$distractors, length) + 1
  }

  p1 <- ggplot(item_metadata, aes(x = factor(num_options))) +
    geom_bar(fill = "#dad7cd") +
    labs(title = "Number of Answer Options by Item", x = "Number of Options", y = "Count") +
    theme_minimal()

  p2 <- ggplot(item_metadata, aes(x = factor(correct_answer, levels = c("A", "B", "C", "D", "E")))) +
    geom_bar(fill = "#ede0bf") +
    labs(title = "Distribution of Correct Answers", x = "Correct Answer", y = "Count") +
    theme_minimal()

  p1 + p2 # Patchwork kombiniert diese zu 1x2 Subplots
}

# 3. Item Parameters Scatter (a vs b)
plot_item_parameters_scatter <- function(item_metadata) {
  ggplot(item_metadata, aes(x = b_param, y = a_param, color = item_type)) +
    geom_point(size = 2, alpha = 0.8) +
    scale_color_manual(values = item_type_colors) +
    labs(title = "Distribution of Item Parameters (a vs b)", 
         x = "Difficulty (b)", y = "Discrimination (a)", color = "Item Type") +
    theme_minimal()
}

# 4. Parameter Distributions (KDE)
plot_parameter_distributions <- function(item_metadata) {
  p1 <- ggplot(item_metadata, aes(x = b_param, fill = item_type)) +
    geom_density( bw = "nrd", trim = FALSE, alpha = 0.25) +
    #geom_density(aes(y = after_stat(density * n / sum(n))), bw = "nrd", trim = FALSE, alpha = 0.25) +
    scale_fill_manual(values = item_type_colors) +
    labs(title = "Difficulty (b)", x = "b") +
    theme_minimal() + theme(legend.position = "none") +
    lims(x=c(-5, 3))

  p2 <- ggplot(item_metadata, aes(x = a_param, fill = item_type)) +
    geom_density(alpha = 0.25, bw = "nrd") +
    scale_fill_manual(values = item_type_colors) +
    labs(title = "Discrimination (a)", x = "a") +
    theme_minimal() +
    lims(x=c(-0.3, 2.5))

  p1 + p2
}

# 5. Slip Distribution (Boxplot)
plot_slip_distribution <- function(item_metadata) {
  ggplot(item_metadata, aes(x = item_type, y = slip, fill = item_type)) +
    geom_boxplot(linewidth = 0.5) +
    scale_fill_manual(values = item_type_colors) +
    labs(title = "Slip Distribution by Item Type", x = "Item Type", y = "Slip") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 20, vjust = 1, hjust = 1),
          legend.position = "none")
}

# 6. Ability Distribution (Theta)
plot_ability_distribution <- function(testtakers) {
  ggplot(testtakers, aes(x = theta_true)) +
    geom_density(fill = "#3B5070", alpha = 0.3) +
    geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.6) +
    labs(title = "Ability Distribution (θ)", x = "θ", y = "Density") +
    theme_minimal()
}

# 7. Score Analysis (Hist + Scatter)
plot_score_analysis <- function(testtakers) {
  mean_val <- mean(testtakers$score_pct)
  med_val <- median(testtakers$score_pct)

  p1 <- ggplot(testtakers, aes(x = score_pct)) +
    geom_histogram(aes(y = after_stat(density)), fill = "#4C72B0", bins = 30, alpha = 0.7) +
    geom_density(color = "black") +
    geom_vline(aes(xintercept = mean_val, color = "Mean"), linetype = "dashed") +
    geom_vline(aes(xintercept = med_val, color = "Median"), linetype = "dotdash") +
    scale_color_manual(name = "Stats", values = c(Mean = "#D62728", Median = "#2CA02C")) +
    labs(title = "Score Distribution", x = "Total Score (%)", y = "Density") +
    theme_minimal()

  p2 <- ggplot(testtakers, aes(x = theta_true, y = score_pct)) +
    geom_point(color = "#4C72B0", alpha = 0.5) +
    labs(title = "Score (%) vs Ability (θ)", x = "Ability (θ)", y = "Total Score (%)") +
    theme_minimal()

  p1 + p2
}

```

```{r plots_print, fig.width=12, fig.height=7}
#| echo: false
#| warning: false

item_metadata = results$item_metadata
testtakers = results$testtakers

# Generate all visualizations
"Generating visualizations...\n"

cat("1. Item Type Distribution")
plot_item_type_distribution(item_metadata)

cat("2. Answer Options Analysis")
plot_answer_options(item_metadata)

cat("3. Item Parameters Scatter Plot")
plot_item_parameters_scatter(item_metadata)

cat("4. Parameter Distributions")
plot_parameter_distributions(item_metadata)

cat("5. Slip Parameter Distribution")
plot_slip_distribution(item_metadata)

cat("6. Ability Distribution")
plot_ability_distribution(testtakers)

cat("7. Score Analysis")
plot_score_analysis(testtakers)

cat("Visualization complete!")
```

```{r results_answer_mat}
#| 
results$answer_matrix
```
