---
title: Synthetic Assessment Data Pipeline (SAD-Pipe) - Analysis
output: html_document
editor_options: 
  markdown: 
    wrap: 120
---

```{r library}
#| include: false

library(tidyverse)
library(ggplot2)
library(purrr)
library(here)
library(rstan)
library(patchwork)
library(knitr)
```

## Visual Encoding Reference

-   This section defines all color palettes and colormaps used throughout the analysis.
-   Centralizing the visual encodings ensures stylistic consistency and accessibility

All palettes have been checked for common forms of color vision deficiency.

```{r colorswatches}
#| echo: false

plot_color_swatch <- function(color_dict, title) {
  df <- data.frame(
    label = names(color_dict),
    color = unlist(color_dict),
    x = seq_along(color_dict)
  )
  
  ggplot(df, aes(x = x, y = 1, fill = color)) +
    geom_tile() +
    scale_fill_identity() +
    geom_text(aes(y = 0.4, label = label), vjust = 1, size = 3.5) +
    labs(title = title) +
    coord_fixed(ratio = 0.5, clip = "off") +
    theme_void() +
    theme(plot.title = element_text(hjust = 0.5, size = 12, margin = margin(b = 10)))
}


# --- Colors for answers (A-E) ---
get_colors <- function(correct_answer = "correct") {
  base_colors <- list(
    'A' = '#e09f3e', 'B' = '#bf6535', 'C' = '#9E2A2B', 
    'D' = '#791B1D', 'E' = '#540b0e', 'correct' = '#ede0bf'
  )

  if (correct_answer %in% names(base_colors)) {
    base_colors[[correct_answer]] <- '#ede0bf'
  }
  return(base_colors)
}

# --- Item-Type Colors ---
item_type_colors <- list(
  'regular'       = '#3B5070',
  'regular_clean' = '#765B7A',
  'diluted'       = '#B56576',
  'misleading'    = '#eaac8b'
)


colors <- get_colors()
plot_color_swatch(colors, "Answer Choice Colors")
plot_color_swatch(item_type_colors, "Item Type Colors")

```

## Load Data

Load the simulated dataset (answer matrix, scored responses, item metadata, and person metadata).

PKL/Rds files are used for fast loading; CSV versions are available if preferred.


```{r datainit}
#| include: false


DATA_DIR_CANDIDATES <- c("data/simulation_results", "data", ".")

DATA_DIR <- NULL
for (d in DATA_DIR_CANDIDATES) {
  potential_path <- here(d)
  if (file.exists(file.path(potential_path, "answer_matrix.Rds"))) {
    DATA_DIR <- potential_path
    break
  }
}

if (is.null(DATA_DIR)) {
  stop(paste(
    "Simulation output files were not found.",
    "One of the following directories must contain answer_matrix.Rds:",
    paste(DATA_DIR_CANDIDATES, collapse = ", "),
    ". Run the simulation notebook first or adjust DATA_DIR_CANDIDATES."
  ))
}

#print(paste("Loaded simulation data from:", fs::path_abs(DATA_DIR)))
answer_matrix <- readRDS(file.path(DATA_DIR, "answer_matrix.Rds"))
scored_responses <- readRDS(file.path(DATA_DIR, "scored_responses.Rds"))
testtakers_post_analysis <- readRDS(file.path(DATA_DIR, "testtakers.Rds"))
item_analysis <- readRDS(file.path(DATA_DIR, "item_metadata.Rds"))

ANALYSIS_DIR <- file.path(fs::path_dir(DATA_DIR), "analysis_results")
fs::dir_create(ANALYSIS_DIR, recurse = TRUE) 

BASE_DIR <- file.path(ANALYSIS_DIR, "item_plots")
fs::dir_create(BASE_DIR, recurse = TRUE)
```

```{r databinary}
#| include: false
#| 
# prepare the data for CTT and IRT
make_binary_matrix <- function(answer_matrix, item_analysis) {
  # Converts the categorical answer matrix into a binary correctness matrix.
  # Each cell is 1 if the examinee selected the correct option, 0 otherwise.
  
  answer_key <- setNames(item_analysis$correct_answer,
                         item_analysis$item)

  items <- names(answer_key)
  
  binary <- map2_dfc(answer_matrix[items], answer_key, ~ as.integer(.x == .y))
  
  return(binary)
}

binary <- make_binary_matrix(answer_matrix, item_analysis)

```

## Classical Test Theory (CTT) Diagnostics

This section computes core CTT statistics to evaluate item functioning:

-   item difficulty
-   item discrimination (item–total correlation, corrected)
-   internal consistency (Cronbach’s alpha)

These diagnostics provide a first check of item quality and help verify that the simulated response patterns align with
the intended psychometric structure.

```{r cttia}
#| echo: false

#' Computes item difficulty and corrected item–total correlations.
#'
#' @param binary A data.frame or matrix of 0/1 correctness 
#'   (rows = persons, columns = items).
#'
#' @return A data.frame with the following columns:
#'   \itemize{
#'     \item \code{item}: Name of the item
#'     \item \code{difficulty}: Item difficulty (p-value)
#'     \item \code{discrimination}: Discrimination (corrected item–total correlation)
#'   }
#'   
classical_item_analysis <- function(binary) {
  total_score <- rowSums(binary)
  
  results <- map_dfr(colnames(binary), function(item) {
    p_value <- mean(binary[[item]], na.rm = TRUE)
    total_score_excl <- total_score - binary[[item]]

    corr <- cor(binary[[item]], total_score_excl, use = "complete.obs")
    tibble(
      item = item,
      difficulty = round(p_value, 3),
      discrimination = round(corr, 3)
    )
  })
  
  return(results)
}

# --- Cronbach's Alpha ---
cronbach_alpha <- function(binary) {
  k <- ncol(binary)
  item_vars <- sapply(binary, var, na.rm = TRUE)
  total_var <- var(rowSums(binary), na.rm = TRUE)
  
  alpha <- (k / (k - 1)) * (1 - sum(item_vars) / total_var)
  return(round(alpha, 3))
}

```

### Run CTT Analysis

```{r cttarung}
#| echo: false
# --- Run CTT Analysis ---
ctt_results = classical_item_analysis(binary)
alpha_value = cronbach_alpha(binary)

head(ctt_results)
```

```{r iactt}
#| include: false
#| 
# --- Merge CTT results with item metadata ---
item_analysis_ctt = left_join(item_analysis,
    ctt_results,
    by="item"
)
```

```{r iacttpic}
#| echo: false
#| 
cat(paste0("Cronbach's alpha: ", alpha_value, ""))

cat("Difficulty summary:")
print(summary(item_analysis_ctt$difficulty))
cat("")

cat("Discrimination (item–total correlation) summary:")
print(summary(item_analysis_ctt$discrimination))
cat("")

# --- Difficulty vs Discrimination Plot (CTT) ---
ggplot(item_analysis_ctt, aes(x = difficulty, 
                              y = discrimination, 
                              color = item_type)) +
  geom_point(size = 2.5, alpha = 0.75) +
  scale_color_manual(values = item_type_colors) +
  labs(
    title = "Item Difficulty vs Discrimination",
    x = "Difficulty (proportion correct)",
    y = "Discrimination (item–total correlation)",
    color = "Item Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold"),
    axis.title = element_text(size = 11),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "right"
  )

```

## Bayesian 2PL IRT model

This section fits a Bayesian 2‑Parameter Logistic (2PL) IRT model to the binary response matrix. The model estimates:

-   **θₚ**: latent ability for each person
-   **aᵢ**: item discrimination (slope of the ICC)
-   **bᵢ**: item difficulty (location on the ability scale)

### Priors

-   **aᵢ \~ LogNormal(0, 0.5)** ensures strictly positive discrimination
-   **bᵢ \~ Normal(0, 2)** provides a weakly informative location prior
-   **θₚ \~ Normal(0, 1)** follows the standard IRT identification convention

### Likelihood

For each person *p* and item *i*:

$p_{pi} = \text{logistic}\!\left(a_i(\theta_p - b_i)\right)$

$u_{pi} \sim \text{Bernoulli}(p_{pi})$

This model allows us to recover the generating item parameters and compare them to the empirical CTT statistics.

```{r stan}
#| warning: false
#| message: false 
#| error: false 
#| results: false

stan_data <- list(
  N = nrow(binary),
  I = ncol(binary),
  y = as.matrix(binary)
)

stan_model_code <- "
  data {
    int<lower=1> N;                // Amount Students
    int<lower=1> I;                // Amount Items
    int<lower=0, upper=1> y[N, I]; // Answermatrix
  }
  parameters {
    vector[N] theta;               
    vector<lower=0>[I] a;          // Discrimination (LogNormal via lower=0)
    vector[I] b;                   // Difficulty
  }
  model {
    theta ~ normal(0, 1);          
    a ~ lognormal(0, 0.5);
    b ~ normal(0, 2);
    
    for (n in 1:N) {
      y[n] ~ bernoulli_logit(a .* (theta[n] - b));
    }
  }
  "


fit <- stan(
  model_code = stan_model_code,
  data = stan_data,
  chains = 4,
  iter = 2000,   # 1000 warm-up + 1000 draws
  warmup = 1000,
  cores = 4,
  seed = 42,
  control = list(adapt_delta = 0.9) 
)

a_hat <- summary(fit, pars = "a")$summary[, "mean"]
b_hat <- summary(fit, pars = "b")$summary[, "mean"]

irt_results <- tibble(
  item = colnames(binary),
  a_hat = a_hat,
  b_hat = b_hat
)

```

```{r irtcttia}
# --- Merge IRT results with item metadata ---
item_analysis_ctt_irt = left_join(item_analysis_ctt, 
                              irt_results,
                              by="item")
```

```{r irtsum}
#| echo: false


cat("Difficulty (b) summary:")
print(summary(item_analysis_ctt_irt$b_hat))
cat("Discrimination (a) summary:")
print(summary(item_analysis_ctt_irt$a_hat))

base_theme <- theme_minimal() + 
  theme(panel.grid.minor = element_blank(),
        axis.line = element_line(color = "black"),
        plot.title = element_text(hjust = 0.5, face = "bold"))

# --- Panel 1: b_hat vs a_hat ---
p1 <- ggplot(item_analysis_ctt_irt, aes(x = b_hat, y = a_hat, color = item_type)) +
  geom_point(size = 3, alpha = 0.75) +
  scale_color_manual(values = item_type_colors) +
  labs(title = "Estimated Difficulty vs Discrimination", x = "b_hat_irt", y = "a_hat_irt") +
  base_theme

# --- Panel 2: a_hat vs a_param (Recovery) ---
p2 <- ggplot(item_analysis_ctt_irt, aes(x = a_param, y = a_hat, color = item_type)) +
  geom_point(size = 3, alpha = 0.75) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.4) + # k-- Linie
  scale_color_manual(values = item_type_colors) +
  labs(title = "Discrimination Recovery", x = "a_param (true)", y = "a_hat (estimated)") +
  base_theme

# --- Panel 3: b_hat vs b_param (Recovery) ---
p3 <- ggplot(item_analysis_ctt_irt, aes(x = b_param, y = b_hat, color = item_type)) +
  geom_point(size = 3, alpha = 0.75) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", alpha = 0.4) +
  scale_color_manual(values = item_type_colors) +
  labs(title = "Difficulty Recovery", x = "b_param (true)", y = "b_hat (estimated)") +
  base_theme

(p1 | p2 | p3) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

```

### Comparing Generating and Estimated IRT Parameters

The dataset used in this notebook was created by simulating item responses from a known set of IRT parameters (a_param,
b_param). These values represent the *generating* item characteristics: the discrimination and difficulty that define
the “true” response process in the simulated population.

In the next step, we fit a Bayesian 2PL IRT model to the simulated data. This model estimates its own set of item
parameters (a_hat, b_hat) based solely on the observed responses. Because the estimation model uses different priors,
identification constraints, and numerical properties than the simulation procedure, the recovered parameters will not
match the generating values exactly.

This comparison is therefore not a test of whether the model reproduces the exact numeric values, but whether it
recovers the *structure* of the items: their relative ordering, their separation by item type, and the overall shape of
the parameter space. Good recovery means that items with higher true difficulty also have higher estimated difficulty,
and similarly for discrimination, even if the absolute values differ.

The following plot visualizes this relationship by showing the generating parameters (circles) and the estimated
parameters (crosses), with arrows indicating the shift from true to estimated values.

### Why true and estimated parameters differ

The estimated IRT parameters do not match the generating parameters exactly, and this is expected for several reasons:

1.  **Person parameters introduce stochastic variation.**\
    Item estimates depend on the distribution of θ, so even with perfect model fit the recovered parameters will deviate
    from the generating values.

2.  **The simulation model and the estimation model are not identical.**\
    The data were generated using our own simulation logic, whereas the estimation uses a Bayesian 2PL model with
    different priors and identification constraints. Differences between true and estimated values therefore reflect
    model structure rather than error.

3.  **Bayesian priors induce shrinkage.**\
    Extreme discrimination or difficulty values are regularized toward the center of the prior distribution, which
    improves stability but reduces exact recovery.

4.  **IRT parameters are identifiable only up to a linear transformation.**\
    The absolute scale of (a, b, θ) is arbitrary; only their relative structure matters. Good recovery therefore means
    preserving the ordering and relationships, not reproducing exact numeric values.

Overall, the model performs well when the estimated parameters preserve the relative ordering and item-type structure of
the generating parameters, even if the absolute values differ.

## CTT–IRT (recovered) Comparison

This section compares empirical CTT statistics with the estimated IRT parameters. Whereas CTT difficulty and
discrimination are sample‑dependent descriptive indices, IRT parameters (a, b) represent model‑based latent item
characteristics.

Comparing both perspectives allows us to evaluate:

-   how well empirical difficulty (proportion correct) aligns with IRT difficulty (b)
-   how well empirical discrimination (item–total correlation) aligns with IRT discrimination (a)

Together, these diagnostics provide a comprehensive view of item functioning across both classical and modern test
theory.

```{r plots1}

comparison_theme <- theme_minimal() +
  theme(
    plot.title = element_text(size = 12, face = "bold", hjust = 0.5),
    axis.title = element_text(size = 11),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    legend.position = "bottom"
  )

# 1. CTT Difficulty vs IRT Difficulty (b)
p_diff <- ggplot(item_analysis_ctt_irt, aes(x = difficulty, y = b_hat, color = item_type)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(values = item_type_colors) +
  labs(
    title = "CTT Difficulty vs IRT Difficulty (b)",
    x = "CTT Difficulty (proportion correct)",
    y = "IRT Difficulty (b)"
  ) +
  comparison_theme

# 2. CTT Discrimination vs IRT Discrimination (a)
p_disc <- ggplot(item_analysis_ctt_irt, aes(x = discrimination, y = a_hat, color = item_type)) +
  geom_point(size = 3, alpha = 0.85) +
  scale_color_manual(values = item_type_colors) +
  labs(
    title = "CTT Discrimination vs IRT Discrimination (a)",
    x = "CTT Discrimination (item–total correlation)",
    y = "IRT Discrimination (a)"
  ) +
  comparison_theme

(p_diff | p_disc) + 
  plot_layout(guides = "collect") & 
  theme(legend.position = "bottom")

```

## Person‑Level Diagnostics

This section documents and checks person-level (testtaker-level) structure.

We analyze relationships between:

-   `theta_true`: generating latent ability (ground truth from the simulation)
-   `theta_post`: posterior mean ability estimate from the Bayesian 2PL model
-   behavioral traits: `carelessness`, `guessiness`, `testwiseness`
-   observed performance: `score`, `score_pct`

These diagnostics serve two purposes:

1.  **Sanity check**: the simulated traits should show interpretable associations with performance.
2.  **Model recovery**: `theta_post` should be strongly associated with `theta_true` (up to estimation error).

Note: correlations involving `score` / `score_pct` are expected to be strong because these are derived from the same
response data used to estimate `theta_post`.

```{r plots2prep}
theta_post_means <- summary(fit, pars = "theta")$summary[, "mean"]
# recovered ability
testtakers_post_analysis <- testtakers_post_analysis %>%
  mutate(theta_post = theta_post_means)
```

```{r plots2}
# Person-level plots + correlations

# Ensure theta_post is present (computed earlier)
if (!"theta_post" %in% colnames(testtakers_post_analysis)) {
  stop("Column `theta_post` missing. Run the Bayesian 2PL IRT model cells first.")
}

cols <- c("theta_true", "theta_post", "carelessness", "guessiness", "testwiseness", "score", "score_pct")

cor_matrix <- testtakers_post_analysis %>%
  select(all_of(cols)) %>%
  cor(use = "complete.obs") %>%
  round(3)

cat("Correlations:")
kable(cor_matrix)

# --- Theta recovery: theta_true vs theta_post ---
p_recovery <- ggplot(testtakers_post_analysis, aes(x = theta_true, y = theta_post)) +
  geom_point(color = "#4C72B0", alpha = 0.6, size = 1.5) +
  labs(title = "Ability Recovery: θ_true vs θ_post",
       x = "θ_true (generating)", y = "θ_post (estimated)") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(), axis.line = element_line(color = "black"))

print(p_recovery)

# --- Traits vs Ability ---
trait_scatter <- function(y_var, title, color) {
  ggplot(testtakers_post_analysis, aes(x = theta_true, y = .data[[y_var]])) +
    geom_point(color = color, alpha = 0.6, size = 1.2) +
    labs(title = title, x = "θ_true", y = y_var) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), axis.line = element_line(color = "black"))
}

p1 <- trait_scatter("carelessness", "Carelessness vs θ_true", "#765B7A")
p2 <- trait_scatter("guessiness", "Guessiness vs θ_true", "#B56576")
p3 <- trait_scatter("testwiseness", "Testwiseness vs θ_true", "#eaac8b")

(p1 | p2 | p3)

# --- 4. Performance vs Ability ---
perf_scatter <- function(x_var, title) {
  ggplot(testtakers_post_analysis, aes(x = .data[[x_var]], y = score_pct)) +
    geom_point(color = "#4C72B0", alpha = 0.6, size = 1.5) +
    labs(title = title, x = x_var, y = "Score (%)") +
    theme_minimal() +
    theme(panel.grid.minor = element_blank(), axis.line = element_line(color = "black"))
}

p4 <- perf_scatter("theta_true", "Score (%) vs θ_true")
p5 <- perf_scatter("theta_post", "Score (%) vs θ_post")

(p4 | p5)

```

## Save the dataframes

```{r savedata}
csv_path <- file.path(ANALYSIS_DIR, "testtakers_post_analysis_r.csv")
rds_path <- file.path(ANALYSIS_DIR, "testtakers_post_analysis.Rds")


write_csv(testtakers_post_analysis, csv_path)

saveRDS(testtakers_post_analysis, rds_path)

message(paste("Daten erfolgreich gespeichert in:", ANALYSIS_DIR))

```

```{r writedata}
write_csv(item_analysis_ctt_irt, file.path(ANALYSIS_DIR, "item_analysis_r.csv"))

saveRDS(item_analysis_ctt_irt, file.path(ANALYSIS_DIR, "item_analysis.Rds"))

```

# **Item‑Level Diagnostics**

To understand how each item behaves across different levels of performance and ability, we generate a **six‑panel
diagnostic grid**. This visualization integrates classical test theory (CTT) and item response theory (IRT), showing how
response patterns relate to both observed scores and model‑based ability estimates.

The six panels provide complementary perspectives:

------------------------------------------------------------------------------------------------------------------------

**1. IRT Item Map (a–b space)**\
Positions the item relative to all others in terms of:

-   **difficulty** (b‑parameter)
-   **discrimination** (a‑parameter)
-   **item type**

This situates the item within the broader structure of the test and highlights whether it behaves similarly to other
items of the same type.

------------------------------------------------------------------------------------------------------------------------

**2. CTT Item Map (difficulty vs. discrimination)**\
Displays the item’s classical difficulty (p‑value) and discrimination (corrected item–total correlation) in the same
way.\
This provides a familiar, score‑based view of item functioning and allows quick comparison with the rest of the test.

------------------------------------------------------------------------------------------------------------------------

**3. IRT Item Characteristic Curve (ICC)**\
Shows the model‑based probability of a correct response as a function of latent ability $\theta$.\
A tangent line at $\theta = b$ visualizes the item’s discrimination (slope), and the b‑parameter is marked explicitly.\
This panel isolates the item’s psychometric behavior from total‑score circularity.

------------------------------------------------------------------------------------------------------------------------

**4. Answer Distribution by Score Group**\
Displays how often each answer option is selected within score‑percentage groups.\
Because these groups are derived from the total score, this panel reflects a **circular relationship**:\
items that contribute strongly to the score naturally show clearer separation between groups.

Useful for:

-   distractor analysis
-   identifying non‑functioning or overly attractive distractors
-   spotting score‑group inconsistencies

------------------------------------------------------------------------------------------------------------------------

**5. Item Summary Block**\
A compact textual summary of key item statistics:

-   CTT difficulty
-   CTT discrimination
-   IRT a‑parameter
-   IRT b‑parameter
-   item type

This provides a quick reference for reviewers.

------------------------------------------------------------------------------------------------------------------------

**6. CTT Discrimination Plot**\
Shows the relationship between:

-   **item score** (0/1)
-   **total score excluding the item**

The red dashed line connects the **group means** for incorrect vs. correct responders, and the corrected item–total
correlation is displayed.\
This panel visualizes how strongly the item separates low‑ and high‑scoring examinees.

------------------------------------------------------------------------------------------------------------------------

**Integrated Interpretation**

Together, these six panels provide a comprehensive view of item behaviour:

-   how distractors function across performance levels
-   how well the ICC fits empirical response patterns
-   how strongly the item discriminates between ability groups
-   whether score‑based and ability‑based patterns diverge
-   how the item compares to the rest of the test in both CTT and IRT space

This integrated view helps identify items that are misfitting, poorly discriminating, too easy or too hard, or
inconsistent across scoring and ability metrics.

```{r plots3prep}
irt_prob <- function(theta, a, b) {
  return(1 / (1 + exp(-a * (theta - b))))
}

plot_item_descriptive_grid <- function(item_analysis, plot_df, binary, item_col) {
  
  meta <- item_analysis %>% filter(item == item_col) %>% slice(1)
  correct_ans_val <- meta$correct_answer
  a <- as.numeric(meta$a_hat)
  b <- as.numeric(meta$b_hat)
  

  current_item_colors <- unlist(get_colors(correct_ans_val))
  type_colors <- unlist(item_type_colors)

  # --- PANEL 1: IRT MAP ---
  p1 <- ggplot(item_analysis, aes(x = a_hat, y = b_hat, color = item_type)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_point(data = meta, size = 5, shape = 21, fill = "transparent", color = "black", stroke = 1.2) +
    scale_color_manual(values = type_colors) +
    labs(title = "IRT item map") + theme_minimal() + theme(legend.position = "none")

  # --- PANEL 2: CTT MAP ---
  p2 <- ggplot(item_analysis, aes(x = discrimination, y = difficulty, color = item_type)) +
    geom_point(alpha = 0.6, size = 2) +
    geom_point(data = meta, size = 5, shape = 21, fill = "transparent", color = "black", stroke = 1.2) +
    scale_color_manual(values = type_colors) +
    labs(title = "CTT item map", color = "Item Type") + theme_minimal()

  # --- PANEL 3: ICC CURVE ---
  theta_range <- seq(-6, 6, length.out = 300)
  icc_data <- data.frame(theta = theta_range, prob = 1 / (1 + exp(-a * (theta_range - b))))
  slope <- a / 4
  
  p3 <- ggplot(icc_data, aes(x = theta, y = prob)) +
    geom_line(linewidth = 1) +
    geom_vline(xintercept = b, linetype = "dashed", color = "grey") +
    annotate(geom = "text", x = b + 0.2, label = paste0("b=", str_sub(as.character(b), 1, 4)), y = 0.15,
             color = "grey", angle = 90, size = 3.5) +
    geom_abline(intercept = 0.5 - slope * b, slope = slope, color = "#297045", linetype = "dotted", linewidth = 1) +
    annotate(geom = "text", x = 5, label = paste0("Tangent (slope=", str_sub(as.character(slope), 1, 4),")"),
             y = 0.05, color = "#297045", size = 2.5) +
    labs(title = paste("IRT curve:", item_col), x = expression(Ability ~ theta), y = "P(correct)") +
    theme_minimal()+
    lims(y = c(0, 1))

  # --- PANEL 4: ANSWER DISTRIBUTION ---
  all_levels <- c(correct_ans_val, meta$distractors[[1]])
  all_labels <- all_levels
  all_labels[1] <- str_c(all_labels[1], "*")
  distr_df <- plot_df %>%
    mutate(score_grp = cut(score/100, breaks = c(-0.01, 0.6, 0.7, 0.8, 0.9, 1.01), 
                           labels = c("0–59%", "60–69%", "70–79%", "80–89%", "90–100%"))) %>%
    group_by(score_grp, ans = !!sym(item_col)) %>%
    summarise(n = n(), .groups = "drop_last") %>%
    mutate(prop = n / sum(n),
           ans = factor(ans, levels = rev(all_levels)))

  count_dist <- distr_df %>%
    group_by(score_grp) %>%
    summarise(n_total = sum(n))
  


  p4 <- ggplot(distr_df, aes(x = score_grp, y = prop, fill = ans)) +
    geom_bar(stat = "identity", position = "stack", color = "white", linewidth = 0.2) +
    geom_text(data = count_dist, 
              aes(x = score_grp, y = 1.02, label = paste0("n=", n_total)), 
              inherit.aes = FALSE,
              size = 3,
              vjust = 0) +
    scale_fill_manual(breaks = all_levels, labels = all_labels, values = current_item_colors, drop = FALSE) +
    scale_y_continuous(limits = c(0, 1.12), labels = scales::percent_format()) +
    labs(title = paste0("Answers (Correct: ", correct_ans_val, ")"),
         x = "Score Group",
         y = "Proportion",
         fill = "Choice") +
    theme_minimal() 

  # --- PANEL 5: SUMMARY TEXT ---
  summary_text <- paste0(
    "Item: ", item_col, "",
    "Type: ", meta$item_type, "",
    "IRT a: ", round(a, 3), "",
    "IRT b: ", round(b, 3), "",
    "Diff (p): ", round(meta$difficulty, 3), "",
    "Discr:  ", round(meta$discrimination, 3)
  )
  p5 <- ggplot() + 
    annotate("text", x = -0.75, y = 2, label = summary_text, family = "mono", hjust = 0.5, vjust = 0.5, size = 3.5) + 
    theme_void()

  # --- PANEL 6: CTT DISCRIMINATION ---
  total_excl <- rowSums(binary[, colnames(binary) != item_col], na.rm = TRUE)
  ctt_df <- data.frame(x = factor(binary[[item_col]], labels = c("Incorrect", "Correct")), y = total_excl)
  
  p6 <- ggplot(ctt_df, aes(x = x, y = y)) +
    geom_jitter(alpha = 0.2, width = 0.15, color = "gray40") +
    stat_summary(fun = mean, geom = "point", size = 4, color = "#297045") +
    stat_summary(fun = mean, geom = "line", aes(group = 1), linetype = "dashed", color = "#297045", linewidth = 1) +
    labs(title = "CTT Discrimination", x = "Item Score", y = "Total (excl. item)") +
    theme_minimal()+
    lims(y = c(0, ncol(plot_df)))

  # --- ASSEMBLY ---
  (p1 | p2) / (p3 | p4) / (p5 | p6) + 
    plot_layout(heights = c(1, 1, 0.7)) & 
    theme(plot.title = element_text(size = 10, face = "bold"))
}

```

```{r plots3}
plot_df <- answer_matrix
plot_df['score'] <-  testtakers_post_analysis['score_pct']
plot_df

plot_item_descriptive_grid(item_analysis_ctt_irt, plot_df, binary, 'item1')
```

### Exporting Item-Level Diagnostic Plots

For transparency and reproducibility, we generate a full diagnostic plot for each item and store the resulting figures
in a dedicated folder. This creates a complete item-level documentation set, allowing reviewers and collaborators to
inspect distractor functioning, score-based and ability-based performance patterns, and the alignment between empirical
data and the estimated IRT model.

```{r plots4}

save_item_figure <- function(plot_obj, base_dir, item, item_type, suffix = "") {
  type_dir <- file.path(base_dir, item_type)
  fs::dir_create(type_dir, recurse = TRUE)
  
  filename <- paste0(item, suffix, ".png")
  filepath <- file.path(type_dir, filename)
  
  ggsave(
    filename = filepath,
    plot = plot_obj,
    dpi = 150,
    width = 14,
    height = 12,
    units = "in",
    device = "png"
  )
}

```

```{r plotswalk}
purrr::walk(item_analysis_ctt_irt$item, function(curr_item) {
  curr_type <- item_analysis_ctt_irt %>% 
    filter(item == curr_item) %>% 
    pull(item_type)

  fig_panel <- plot_item_descriptive_grid(item_analysis_ctt_irt, plot_df, binary, curr_item)
  
  save_item_figure(fig_panel, BASE_DIR, curr_item, curr_type, suffix = "_panel_r")
})

cat("All item plots generated and saved.")
```

```{r}
item_analysis_ctt_irt

```
